---
title: "Optimization in R"
output: html_document
---
Resource: http://www.is.uni-freiburg.de/resources/computational-economics/5_OptimizationR.pdf

```{r One Dimensional Optimization}

# SYNTAX: optimizer(objective, constraints, bounds=NULL, types=NULL, maximum=FALSE)

# One Dimensional Function (Objective Function)
f = function(x) 2*(x[1]-1)^2 + 5*(x[2]-3)^2 + 10

# Constraints
cons = c(1,1)

# Optimization
r = optim(cons, f)
r

# Check if optimization converged to 0
r$convergence == 0

# Input arguments for the optimization function
r$par

# Value of the objective function
r$value

```


# Linear Programming Optimization
Objective function
  - goal is to maximize profits
  - products A and B are sold at 25 and 20
  $\text { Sales_ }_{\max }=\max _{x_{1}, x_{2}} 25 x_{1}+20 x_{2}=\max _{x_{1}, x_{2}}\left[\begin{array}{c}{25} \\ {20}\end{array}\right]^{T}\left[\begin{array}{c}{x_{1}} \\ {x_{2}}\end{array}\right]$
Constraints
  Resource:
  - product A requires 20 units, B requires 12
  - only 1800 untis
  Time:
  - both products require 1/15 hours
  - a working day has 8 hours
  $\begin{array}{l}{20 x_{1}+12 x_{2} \leq 1800} \\ {\frac{1}{15} x_{1}+\frac{1}{15} x_{2} \leq 8}\end{array}$

```{r Linear Programming Optimization}
require(lpSolve)

# SYNTAX: lp(direction="min", objective.in, const.mat, const.dir, const.rhs)

# Solving for where the two lines meet will minimize the objective
objective.in = c(25, 20)
const.mat = matrix(c(20, 12, 1/15, 1/15), nrow=2, byrow=T)
const.rhs = c(1800, 8)
const.dir = c("<=", "<=")
optimum = lp(direction="max", objective.in, const.mat, const.dir, const.rhs)
optimum

# Optimal values of x1 and x2
optimum$solution

# Objective at minimum 
optimum$objval

```



```{r Non-differentiable function with optimx}
f = function(x) return(abs(x-2) + 2*abs(x-1))
xmin = optimize(f, interval = c(0, 3), tol = 0.0001) 
xmin

plot(f, 0, 3)
```


